---
title: "Preliminary_genetic_analyses"
author: "Lior Abramson"
date: "`r Sys.Date()`"
output: html_document
---

This script performs all the preliminary analyses and examines twins' ICC in the study main variables (empathy and empathic personality profiles) before conducting 
the biometric cross-lagged analysis in the Mx software. 
It also calculates the correlations between the study's main variables 
and non-self-report measures of empathy

*'pred'/'predicted' indicates the predicted scores of empathy, i.e., the empathic personality profile scores

Preparations:
```{r}

#Preparations
  rm(list = ls()) # clean the global environment
  cat ("\014")    #clean the R console

#Packages
  packages <- c('broom','purrr','lsr','psych', 'apaTables', 'reshape','irr')
  lapply(packages, require, character.only = TRUE)

#Import the file (D was created using 'the_empathic_personality_profiles.R' script)
  D <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Dfinal_11_13.csv")

```

examining attrition effects: are there differences in the study variables 
between adolescents who participated in one or two time points?
```{r}
#build the variable indicating whether the participant was in one or two time points
#0= one time-point, 1=two time-points
  D$whatStage <- 0
  D$whatStage[is.na(D$EMPQ_emotional_11)== F &is.na(D$EMPQ_emotional_13)== F] <-1    

#perform separate t-test for each study outcome
  ttest_whatStage_emo11 <- t.test(EMPQ_emotional_11~whatStage, data=D)
  ttest_whatStage_cog11 <- t.test(EMPQ_cognitive_11~whatStage, data=D)
  ttest_whatStage_predemo11 <- t.test(predicted_emotional_11~whatStage, data=D)
  ttest_whatStage_predcog11 <- t.test(predicted_cognitive_11~whatStage, data=D)
  
  ttest_whatStage_emo13 <- t.test(EMPQ_emotional_13~whatStage, data=D)
  ttest_whatStage_cog13 <- t.test(EMPQ_cognitive_13~whatStage, data=D)
  ttest_whatStage_predemo13 <- t.test(predicted_emotional_13~whatStage, data=D)
  ttest_whatStage_predcog13 <- t.test(predicted_cognitive_13~whatStage, data=D)

#enter all tests in one table
  ttest_attritionTable <- map_df(list(ttest_whatStage_predemo11,ttest_whatStage_emo11,
                                      ttest_whatStage_predemo13,ttest_whatStage_emo13,
                                      ttest_whatStage_predcog11,ttest_whatStage_cog11,
                                      ttest_whatStage_predcog13,ttest_whatStage_cog13), tidy)

  ttest_attritionTable_edit <- ttest_attritionTable[c("estimate1", "estimate2","statistic","parameter","p.value")]
  colnames(ttest_attritionTable_edit) <- c("Mean one time-point","Mean both time-point", "t value","df","p-value")
  ttest_attritionTable_edit[,1:4] <- round(ttest_attritionTable_edit[,1:4],2)
  ttest_attritionTable_edit[,5]   <- round(ttest_attritionTable_edit[,5],3)

#add effect size
#change the direction according to the t-test results when relevant 
#because the cohensD function does not show the effect's direction 
  ttest_attritionTable_edit$cohenD <- NA
  
  ttest_attritionTable_edit$cohenD[1] <- round(cohensD(predicted_emotional_11~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[2] <- round(cohensD(EMPQ_emotional_11~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[3] <- 0-round(cohensD(predicted_emotional_13~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[4] <- 0-round(cohensD(EMPQ_emotional_13~whatStage, data=D),2)
  
  ttest_attritionTable_edit$cohenD[5] <- 0-round(cohensD(predicted_cognitive_11~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[6] <- 0-round(cohensD(EMPQ_cognitive_11~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[7] <- 0-round(cohensD(predicted_cognitive_13~whatStage, data=D),2)
  ttest_attritionTable_edit$cohenD[8] <- 0-round(cohensD(EMPQ_cognitive_13~whatStage, data=D),2)

#add sd
  ttest_attritionTable_edit$sd1Stage <- NA
  ttest_attritionTable_edit$sd2Stage <- NA
  
  ttest_attritionTable_edit$sd1Stage[1] <- round((aggregate(D$predicted_emotional_11~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[2] <- round((aggregate(D$EMPQ_emotional_11~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[3] <- round((aggregate(D$predicted_emotional_13~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[4] <- round((aggregate(D$EMPQ_emotional_13~D$whatStage, FUN = sd)[1,2]),2)

  ttest_attritionTable_edit$sd1Stage[5] <- round((aggregate(D$predicted_cognitive_11~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[6] <- round((aggregate(D$EMPQ_cognitive_11~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[7] <- round((aggregate(D$predicted_cognitive_13~D$whatStage, FUN = sd)[1,2]),2)
  ttest_attritionTable_edit$sd1Stage[8] <- round((aggregate(D$EMPQ_cognitive_13~D$whatStage, FUN = sd)[1,2]),2)
  
  ttest_attritionTable_edit$sd2Stage[1] <- round((aggregate(D$predicted_emotional_11~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[2] <- round((aggregate(D$EMPQ_emotional_11~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[3] <- round((aggregate(D$predicted_emotional_13~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[4] <- round((aggregate(D$EMPQ_emotional_13~D$whatStage, FUN = sd)[2,2]),2)
  
  ttest_attritionTable_edit$sd2Stage[5] <- round((aggregate(D$predicted_cognitive_11~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[6] <- round((aggregate(D$EMPQ_cognitive_11~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[7] <- round((aggregate(D$predicted_cognitive_13~D$whatStage, FUN = sd)[2,2]),2)
  ttest_attritionTable_edit$sd2Stage[8] <- round((aggregate(D$EMPQ_cognitive_13~D$whatStage, FUN = sd)[2,2]),2)

#adjust labels
  ttest_attritionTable_edit$`Mean one time-point` <- paste(as.character(ttest_attritionTable_edit$`Mean one time-point`),"(",
                                                           as.character (ttest_attritionTable_edit$sd1Stage),")")
  ttest_attritionTable_edit$`Mean both time-point` <- paste(as.character(ttest_attritionTable_edit$`Mean both time-point`),"(",
                                                            as.character (ttest_attritionTable_edit$sd2Stage),")")
  ttest_attritionTable_edit$`t value` <- paste(as.character(ttest_attritionTable_edit$`t value`),"(",
                                               as.character (ttest_attritionTable_edit$df),")")
  
  df_attrition <- apply(ttest_attritionTable_edit[,c(1,2,3,6)],2,as.character)

#export table to csv
  write.csv(df_attrition, file= "ttest_attritionTable.csv", row.names = F)

```

examining relations between sex and the study variables
```{r}

#perform separate t-test for each study outcome
  ttest_sexPredemo11 <- t.test(predicted_emotional_11~sex, data=D)
  ttest_sexEmpemo11 <-t.test(EMPQ_emotional_11~sex, data=D)
  ttest_sexPredcog11 <- t.test(predicted_cognitive_11~sex, data=D)
  ttest_sexEmpcog11 <- t.test(EMPQ_cognitive_11~sex, data=D)
  
  ttest_sexPredemo13 <- t.test(predicted_emotional_13~sex, data=D)
  ttest_sexEmpemo13 <-t.test(EMPQ_emotional_13~sex, data=D)
  ttest_sexPredcog13 <- t.test(predicted_cognitive_13~sex, data=D)
  ttest_sexEmpcog13 <- t.test(EMPQ_cognitive_13~sex, data=D)

#enter all tests in one table
  ttest_sexTable <- map_df(list(ttest_sexPredemo11,ttest_sexEmpemo11,
                                ttest_sexPredemo13,ttest_sexEmpemo13,
                                ttest_sexPredcog11, ttest_sexEmpcog11,
                                ttest_sexPredcog13, ttest_sexEmpcog13), tidy)
  
  ttest_sexTable_edit <- ttest_sexTable[c("estimate1", "estimate2","statistic","parameter","p.value")]
  colnames(ttest_sexTable_edit) <- c("Mean males","Mean females", "t value","df","p-value")
  ttest_sexTable_edit[,1:4] <- round(ttest_sexTable_edit[,1:4],2)
  ttest_sexTable_edit[,5] <- round(ttest_sexTable_edit[,5],3)

#add effect size
#change the direction according to the t-test results when relevant 
#because the cohensD function does not show the effect's direction 
  ttest_sexTable_edit$cohenD <- NA
  
  ttest_sexTable_edit$cohenD[1] <- 0-round(cohensD(predicted_emotional_11~sex, data=D),2)
  ttest_sexTable_edit$cohenD[2] <- 0-round(cohensD(EMPQ_emotional_11~sex, data=D),2)
  ttest_sexTable_edit$cohenD[3] <- 0-round(cohensD(predicted_emotional_13~sex, data=D),2)
  ttest_sexTable_edit$cohenD[4] <- 0-round(cohensD(EMPQ_emotional_13~sex, data=D),2)
  
  ttest_sexTable_edit$cohenD[5] <- 0-round(cohensD(predicted_cognitive_11~sex, data=D),2)
  ttest_sexTable_edit$cohenD[6] <- 0-round(cohensD(EMPQ_cognitive_11~sex, data=D),2)
  ttest_sexTable_edit$cohenD[7] <- 0-round(cohensD(predicted_cognitive_13~sex, data=D),2)
  ttest_sexTable_edit$cohenD[8] <- 0-round(cohensD(EMPQ_cognitive_13~sex, data=D),2)

#add sd
  ttest_sexTable_edit$sdBoys <- NA
  ttest_sexTable_edit$sdGirls <- NA
  
  ttest_sexTable_edit$sdBoys[1] <- round((aggregate(D$predicted_emotional_11~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[2] <- round((aggregate(D$EMPQ_emotional_11~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[3] <- round((aggregate(D$predicted_emotional_13~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[4] <- round((aggregate(D$EMPQ_emotional_13~D$sex, FUN = sd)[1,2]),2)
  
  ttest_sexTable_edit$sdBoys[5] <- round((aggregate(D$predicted_cognitive_11~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[6] <- round((aggregate(D$EMPQ_cognitive_11~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[7] <- round((aggregate(D$predicted_cognitive_13~D$sex, FUN = sd)[1,2]),2)
  ttest_sexTable_edit$sdBoys[8] <- round((aggregate(D$EMPQ_cognitive_13~D$sex, FUN = sd)[1,2]),2)
  
  ttest_sexTable_edit$sdGirls[1] <- round((aggregate(D$predicted_emotional_11~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[2] <- round((aggregate(D$EMPQ_emotional_11~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[3] <- round((aggregate(D$predicted_emotional_13~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[4] <- round((aggregate(D$EMPQ_emotional_13~D$sex, FUN = sd)[2,2]),2)
  
  ttest_sexTable_edit$sdGirls[5] <- round((aggregate(D$predicted_cognitive_11~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[6] <- round((aggregate(D$EMPQ_cognitive_11~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[7] <- round((aggregate(D$predicted_cognitive_13~D$sex, FUN = sd)[2,2]),2)
  ttest_sexTable_edit$sdGirls[8] <- round((aggregate(D$EMPQ_cognitive_13~D$sex, FUN = sd)[2,2]),2)
  
#edit labels
  ttest_sexTable_edit$`Mean males` <- paste(as.character(ttest_sexTable_edit$`Mean males`),"(",
                                           as.character (ttest_sexTable_edit$sdBoys),")")
  ttest_sexTable_edit$`Mean females` <- paste(as.character(ttest_sexTable_edit$`Mean females`),"(",
                                           as.character (ttest_sexTable_edit$sdGirls),")")
  ttest_sexTable_edit$`t value` <- paste(as.character(ttest_sexTable_edit$`t value`),"(",
                                         as.character (ttest_sexTable_edit$df),")")
#add stars indicating significance
  for (i in 1:8) { ifelse(ttest_sexTable_edit$`p-value`[i]<.0001,
        ttest_sexTable_edit$`t value`[i]<- paste0 (ttest_sexTable_edit$`t value`[i],"***"),
          ifelse(ttest_sexTable_edit$`p-value`[i]<.01,
                 ttest_sexTable_edit$`t value`[i]<- paste0 (ttest_sexTable_edit$`t value`[i],"**"), 
                    ttest_sexTable_edit$`t value`[i]<- ttest_sexTable_edit$`t value`[i]))}
  
  df_sex <- apply(ttest_sexTable_edit[,c(1,2,3,6)],2,as.character)

#export table to csv
  write.csv(df_sex, file= "ttest_sex.csv", row.names = F)

```

examining relations between questionnaire format (online/pencil and paper) and the study variables
```{r}

# at age 13 time-point, some children received a shorter battery that did not include 
# the emotion recognition test to increase participation rate.
# as a first step, conduct t-test only for the difference between the short and long online versions:
  t.test(predicted_emotional_13~OnlineManual13, data=D[D$OnlineManual13<=1,])
  t.test(EMPQ_emotional_13~OnlineManual13, data=D[D$OnlineManual13<=1,])
  t.test(predicted_cognitive_13~OnlineManual13, data=D[D$OnlineManual13<=1,])
  t.test(EMPQ_cognitive_13~OnlineManual13, data=D[D$OnlineManual13<=1,])

#because there is no significant difference, unite the two online groups
  D$OnlineManual13_2G <- D$OnlineManual13
  D$OnlineManual13_2G[D$OnlineManual13==0] <-"online"
  D$OnlineManual13_2G[D$OnlineManual13==1] <-"online"
  D$OnlineManual13_2G[D$OnlineManual13==2] <-"paper"

#perform separate t-test for each study outcome
  ttest_sourcePredemo11 <- t.test(predicted_emotional_11~OnlineManual, data=D)
  ttest_sourceEmpemo11 <-t.test(EMPQ_emotional_11~ OnlineManual, data=D)
  ttest_sourcePredcog11 <- t.test(predicted_cognitive_11~ OnlineManual, data=D)
  ttest_sourceEmpcog11 <- t.test(EMPQ_cognitive_11~ OnlineManual, data=D)
  
  ttest_sourcePredemo13 <- t.test(predicted_emotional_13~ OnlineManual13_2G, data=D)
  ttest_sourceEmpemo13 <-t.test(EMPQ_emotional_13~ OnlineManual13_2G, data=D)
  ttest_sourcePredcog13 <- t.test(predicted_cognitive_13~ OnlineManual13_2G, data=D)
  ttest_sourceEmpcog13 <- t.test(EMPQ_cognitive_13~ OnlineManual13_2G, data=D)
  
#enter all tests in one table
  ttest_sourceTable <- map_df(list(ttest_sourcePredemo11,ttest_sourceEmpemo11,
                                   ttest_sourcePredemo13,ttest_sourceEmpemo13,
                                   ttest_sourcePredcog11, ttest_sourceEmpcog11,
                                   ttest_sourcePredcog13, ttest_sourceEmpcog13), tidy)
  
  ttest_sourceTable_edit <- ttest_sourceTable[c("estimate1", "estimate2","statistic","parameter","p.value")]
  colnames(ttest_sourceTable_edit) <- c("MeanOnline","MeanPaper", "t_value","df","p_value")
  ttest_sourceTable_edit[,1:4] <- round(ttest_sourceTable_edit[,1:4],2)
  ttest_sourceTable_edit[,5] <- round(ttest_sourceTable_edit[,5],3)


#add effect size
#change the direction according to the t-test results when relevant 
#because the cohensD function does not show the effect's direction 
  ttest_sourceTable_edit$cohenD <- NA
  
  ttest_sourceTable_edit$cohenD[1] <- 0-round(cohensD(predicted_emotional_11~OnlineManual, data=D),2)
  ttest_sourceTable_edit$cohenD[2] <- 0-round(cohensD(EMPQ_emotional_11~OnlineManual, data=D),2)
  ttest_sourceTable_edit$cohenD[3] <- round(cohensD(predicted_emotional_13~OnlineManual13_2G, data=D),2)
  ttest_sourceTable_edit$cohenD[4] <- 0-round(cohensD(EMPQ_emotional_13~OnlineManual13_2G, data=D),2)
  
  ttest_sourceTable_edit$cohenD[5] <- 0-round(cohensD(predicted_cognitive_11~OnlineManual, data=D),2)
  ttest_sourceTable_edit$cohenD[6] <- 0-round(cohensD(EMPQ_cognitive_11~OnlineManual, data=D),2)
  ttest_sourceTable_edit$cohenD[7] <- round(cohensD(predicted_cognitive_13~OnlineManual13_2G, data=D),2)
  ttest_sourceTable_edit$cohenD[8] <- round(cohensD(EMPQ_cognitive_13~OnlineManual13_2G, data=D),2)
  

#add sd
  ttest_sourceTable_edit$sdOnline <- NA
  ttest_sourceTable_edit$sdPaper <- NA
  
  ttest_sourceTable_edit$sdOnline[1] <- round((aggregate(D$predicted_emotional_11~D$OnlineManual, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[2] <- round((aggregate(D$EMPQ_emotional_11~D$OnlineManual, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[3] <- round((aggregate(D$predicted_emotional_13~D$OnlineManual13_2G, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[4] <- round((aggregate(D$EMPQ_emotional_13~D$OnlineManual13_2G, FUN = sd)[1,2]),2)
  
  ttest_sourceTable_edit$sdOnline[5] <- round((aggregate(D$predicted_cognitive_11~D$OnlineManual, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[6] <- round((aggregate(D$EMPQ_cognitive_11~D$OnlineManual, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[7] <- round((aggregate(D$predicted_cognitive_13~D$OnlineManual13_2G, FUN = sd)[1,2]),2)
  ttest_sourceTable_edit$sdOnline[8] <- round((aggregate(D$EMPQ_cognitive_13~D$OnlineManual13_2G, FUN = sd)[1,2]),2)
  
  ttest_sourceTable_edit$sdPaper[1] <- round((aggregate(D$predicted_emotional_11~D$OnlineManual, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[2] <- round((aggregate(D$EMPQ_emotional_11~D$OnlineManual, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[3] <- round((aggregate(D$predicted_emotional_13~D$OnlineManual13_2G, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[4] <- round((aggregate(D$EMPQ_emotional_13~D$OnlineManual13_2G, FUN = sd)[2,2]),2)
  
  ttest_sourceTable_edit$sdPaper[5] <- round((aggregate(D$predicted_cognitive_11~D$OnlineManual, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[6] <- round((aggregate(D$EMPQ_cognitive_11~D$OnlineManual, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[7] <- round((aggregate(D$predicted_cognitive_13~D$OnlineManual13_2G, FUN = sd)[2,2]),2)
  ttest_sourceTable_edit$sdPaper[8] <- round((aggregate(D$EMPQ_cognitive_13~D$OnlineManual13_2G, FUN = sd)[2,2]),2)

#edit labels
  ttest_sourceTable_edit$MeanOnline <- paste(as.character(ttest_sourceTable_edit$MeanOnline),"(",
                                           as.character (ttest_sourceTable_edit$sdOnline),")")
  ttest_sourceTable_edit$MeanPaper <- paste(as.character(ttest_sourceTable_edit$MeanPaper),"(",
                                            as.character (ttest_sourceTable_edit$sdPaper),")")
  ttest_sourceTable_edit$t_value <- paste(as.character(ttest_sourceTable_edit$t_value),"(",
                                         as.character (ttest_sourceTable_edit$df),")")

#add stars that indicate significance
  for (i in 1:8) { ifelse(ttest_sourceTable_edit$p_value[i]<.001,
                          ttest_sourceTable_edit$t_value[i]<- paste0 (ttest_sourceTable_edit$t_value[i],"***"),
                          ifelse(ttest_sourceTable_edit$p_value[i]<.01,
                                 ttest_sourceTable_edit$t_value[i]<- paste0 (ttest_sourceTable_edit$t_value[i],"**"),
                                 ifelse(ttest_sourceTable_edit$p_value[i]<.05,
                                        ttest_sourceTable_edit$t_value[i]<- paste0 (ttest_sourceTable_edit$t_value[i],"*"),
                                 ttest_sourceTable_edit$t_value[i]<- ttest_sourceTable_edit$t_value[i])))}
  
  df_source <- apply(ttest_sourceTable_edit[,c(1,2,3,6)],2,as.character)

#export table to csv
  write.csv(df_source, file= "ttest_source.csv", row.names = F)


```

examining correlations between study variables and age within each time-point
```{r}

#import participants age in each wave
  participants_ages11 <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Age11_participants_age.csv")
  participants_ages13 <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Age13_participants_age.csv")
  
  D<- merge(x=D, y=participants_ages11, by=c("ifam","ID"), all.x=T, all.y=F )     
  D<- merge(x=D, y=participants_ages13, by=c("ifam","ID"), all.x=T, all.y=F ) 

#examine the correlations
  col=colnames(D)
  relvar_agecor11 <- c(which (col=="Age11"),
                       which (col=="predicted_emotional_11"),which(col=="predicted_cognitive_11"),
                       which (col=="EMPQ_emotional_11"),which(col=="EMPQ_cognitive_11"))
                       
  apa.cor.table(D[,relvar_agecor11])

  relvar_agecor13 <- c(which (col=="Age13"),
                     which (col=="predicted_emotional_13"),which(col=="predicted_cognitive_13"),
                     which (col=="EMPQ_emotional_13"),which(col=="EMPQ_cognitive_13"))

  apa.cor.table(D[,relvar_agecor13])

```
standardization according to sex-saving residuals cleaned from sex
```{r}
#creating a matrix to store the data
  col <- colnames(D)
  relvar <- c(which (col=="EMPQ_emotional_11"),which(col=="EMPQ_cognitive_11"),
              which(col=="predicted_emotional_11"), which(col=="predicted_cognitive_11"),
              which (col=="EMPQ_emotional_13"),which(col=="EMPQ_cognitive_13"),
              which(col=="predicted_emotional_13"), which(col=="predicted_cognitive_13"))

  residvec <- paste0(colnames(D[,relvar]),"r")
  resid.matrix <- as.data.frame(matrix(data= NA, nrow =nrow(D),ncol =length(relvar)))
  colnames(resid.matrix) <-residvec

#creating the residual scores
  for (i in 1:length(relvar)) {
        reg <- lm(D[,relvar[i]]~sex, data=D, na.action=na.exclude)
        resid.matrix[,i] <-resid(reg)
        }
  D <-cbind(D,resid.matrix)

```


examining differences between DZ same-sex (DZs) and DZ opposite sex (DZo) twins
in order to decide whether indeed these groups can be united
```{r}
#Converting dataset to wide format
  DWide <- reshape(D, timevar = "ID", idvar = "ifam", direction = "wide", 
                        v.names = colnames(D[-c(1,2,which(colnames(D)=="zygosity"))]))
   
  newnames <- gsub(x=colnames(DWide),pattern="[.]", replacement="_")
  newnames <- gsub(x=newnames,pattern="_4", replacement="_2")
  newnames <- gsub(x=newnames,pattern="_11", replacement="11")
  newnames <- gsub(x=newnames,pattern="_13", replacement="13")
  colnames(DWide) <- newnames

#create different datasets for the 3 zygosity groups 
  Dmz  = subset(DWide, zygosity == 1)
  Ddzs = subset(DWide, zygosity == 2)
  Ddzo = subset(DWide, zygosity == 3)

#computing ICC between twins separately for each zygosity group
  col <- colnames(DWide)
  
  #age 11
  icc_EMPemo_MZ11  <- icc(Dmz[,c(which(col=="EMPQ_emotional11_1"),which(col=="EMPQ_emotional11_2"))],model="oneway", unit="single")
  icc_EMPemo_DZS11 <- icc(Ddzs[,c(which(col=="EMPQ_emotional11_1"),which(col=="EMPQ_emotional11_2"))],model="oneway", unit="single")
  icc_EMPemo_DZO11 <- icc(Ddzo[,c(which(col=="EMPQ_emotional11_1"),which(col=="EMPQ_emotional11_2"))],model="oneway", unit="single")
  
  icc_EMPcog_MZ11  <- icc(Dmz[,c(which(col=="EMPQ_cognitive11_1"),which(col=="EMPQ_cognitive11_2"))],model="oneway", unit="single")
  icc_EMPcog_DZS11 <-icc(Ddzs[,c(which(col=="EMPQ_cognitive11_1"),which(col=="EMPQ_cognitive11_2"))],model="oneway", unit="single")
  icc_EMPcog_DZO11 <-icc(Ddzo[,c(which(col=="EMPQ_cognitive11_1"),which(col=="EMPQ_cognitive11_2"))],model="oneway", unit="single")
  
  icc_PREDemo_MZ11 <-icc(Dmz[,c(which(col=="predicted_emotional11_1"),which(col=="predicted_emotional11_2"))],model="oneway", unit="single")
  icc_PREDemo_DZS11 <-icc(Ddzs[,c(which(col=="predicted_emotional11_1"),which(col=="predicted_emotional11_2"))],model="oneway", unit="single")
  icc_PREDemo_DZO11 <-icc(Ddzo[,c(which(col=="predicted_emotional11_1"),which(col=="predicted_emotional11_2"))],model="oneway", unit="single")
  
  icc_PREDcog_MZ11  <- icc(Dmz[,c(which(col=="predicted_cognitive11_1"),which(col=="predicted_cognitive11_2"))],model="oneway", unit="single")
  icc_PREDcog_DZS11 <- icc(Ddzs[,c(which(col=="predicted_cognitive11_1"),which(col=="predicted_cognitive11_2"))],model="oneway", unit="single")
  icc_PREDcog_DZO11 <- icc(Ddzo[,c(which(col=="predicted_cognitive11_1"),which(col=="predicted_cognitive11_2"))],model="oneway", unit="single")

  #fisher Z comparison between DZs and DZo ICC
  fishZ_empEmo11  <- r.test(n=icc_EMPemo_DZS11$subjects, n2=icc_EMPemo_DZO11$subjects,r12=icc_EMPemo_DZS11$value,  r34=icc_EMPemo_DZO11$value)  
  fishZ_empCog11  <- r.test(n=icc_EMPcog_DZS11$subjects, n2=icc_EMPcog_DZO11$subjects,r12=icc_EMPcog_DZS11$value,  r34=icc_EMPcog_DZO11$value)     
  fishZ_predEmo11 <- r.test(n=icc_PREDemo_DZS11$subjects, n2=icc_PREDemo_DZO11$subjects,r12=icc_PREDemo_DZS11$value,  r34=icc_PREDemo_DZO11$value) 
  fishZ_predCog11 <- r.test(n=icc_PREDcog_DZS11$subjects, n2=icc_PREDcog_DZO11$subjects,r12=icc_PREDcog_DZS11$value,  r34=icc_PREDcog_DZO11$value) 

  #age 13
  icc_EMPemo_MZ13  <- icc(Dmz[,c(which(col=="EMPQ_emotional13_1"),which(col=="EMPQ_emotional13_2"))],model="oneway", unit="single")
  icc_EMPemo_DZS13 <- icc(Ddzs[,c(which(col=="EMPQ_emotional13_1"),which(col=="EMPQ_emotional13_2"))],model="oneway", unit="single")
  icc_EMPemo_DZO13 <- icc(Ddzo[,c(which(col=="EMPQ_emotional13_1"),which(col=="EMPQ_emotional13_2"))],model="oneway", unit="single")
  
  icc_EMPcog_MZ13  <- icc(Dmz[,c(which(col=="EMPQ_cognitive13_1"),which(col=="EMPQ_cognitive13_2"))],model="oneway", unit="single")
  icc_EMPcog_DZS13 <-icc(Ddzs[,c(which(col=="EMPQ_cognitive13_1"),which(col=="EMPQ_cognitive13_2"))],model="oneway", unit="single")
  icc_EMPcog_DZO13 <-icc(Ddzo[,c(which(col=="EMPQ_cognitive13_1"),which(col=="EMPQ_cognitive13_2"))],model="oneway", unit="single")
  
  icc_PREDemo_MZ13 <-icc(Dmz[,c(which(col=="predicted_emotional13_1"),which(col=="predicted_emotional13_2"))],model="oneway", unit="single")
  icc_PREDemo_DZS13 <-icc(Ddzs[,c(which(col=="predicted_emotional13_1"),which(col=="predicted_emotional13_2"))],model="oneway", unit="single")
  icc_PREDemo_DZO13 <-icc(Ddzo[,c(which(col=="predicted_emotional13_1"),which(col=="predicted_emotional13_2"))],model="oneway", unit="single")
  
  icc_PREDcog_MZ13  <- icc(Dmz[,c(which(col=="predicted_cognitive13_1"),which(col=="predicted_cognitive13_2"))],model="oneway", unit="single")
  icc_PREDcog_DZS13 <- icc(Ddzs[,c(which(col=="predicted_cognitive13_1"),which(col=="predicted_cognitive13_2"))],model="oneway", unit="single")
  icc_PREDcog_DZO13 <- icc(Ddzo[,c(which(col=="predicted_cognitive13_1"),which(col=="predicted_cognitive13_2"))],model="oneway", unit="single")

  #fisher Z comparison between DZS and DZo ICC
  fishZ_empEmo13  <- r.test(n=icc_EMPemo_DZS13$subjects, n2=icc_EMPemo_DZO13$subjects,r12=icc_EMPemo_DZS13$value,  r34=icc_EMPemo_DZO13$value) 
  fishZ_empCog13  <- r.test(n=icc_EMPcog_DZS13$subjects, n2=icc_EMPcog_DZO13$subjects,r12=icc_EMPcog_DZS13$value,  r34=icc_EMPcog_DZO13$value) 
  fishZ_predEmo13 <- r.test(n=icc_PREDemo_DZS13$subjects, n2=icc_PREDemo_DZO13$subjects,r12=icc_PREDemo_DZS13$value,  r34=icc_PREDemo_DZO13$value) 
  fishZ_predCog13 <- r.test(n=icc_PREDcog_DZS13$subjects, n2=icc_PREDcog_DZO13$subjects,r12=icc_PREDcog_DZS13$value,  r34=icc_PREDcog_DZO13$value) 
  
  #planting Fisher Z results in a table
  #when relevant, the direction of the Z-value was manually changed in the word document, not here
  #when relevant, stars indicating significance were manually entered in the word document
  fishZ_table <- matrix(nrow=8,ncol=4)
  fishZ_table[,1] <- c("Emotional empathic personality profile 11","Emotional empathy 11",
                       "Emotional empathic personality profile 13","Emotional empathy 13",
                       "Cognitive empathic personality profile 11","Cognitive empathy 11",
                       "Cognitive empathic personality profile 13","Cognitive empathy 13")
  
  fishZ_table[1,2] <- paste(as.character(round(icc_PREDemo_DZS11$value,2)),"[",
                            as.character(round(icc_PREDemo_DZS11$lbound,2)),"-",
                            as.character(round(icc_PREDemo_DZS11$ubound,2)),"]") 

  fishZ_table[2,2] <- paste(as.character(round(icc_EMPemo_DZS11$value,2)),"[",
                            as.character(round(icc_EMPemo_DZS11$lbound,2)),"-",
                            as.character(round(icc_EMPemo_DZS11$ubound,2)),"]")
  
  fishZ_table[3,2] <- paste(as.character(round(icc_PREDemo_DZS13$value,2)),"[",
                            as.character(round(icc_PREDemo_DZS13$lbound,2)),"-",
                            as.character(round(icc_PREDemo_DZS13$ubound,2)),"]") 
  
  fishZ_table[4,2] <- paste(as.character(round(icc_EMPemo_DZS13$value,2)),"[",
                            as.character(round(icc_EMPemo_DZS13$lbound,2)),"-",
                            as.character(round(icc_EMPemo_DZS13$ubound,2)),"]")
  
  
  fishZ_table[5,2] <- paste(as.character(round(icc_PREDcog_DZS11$value,2)),"[",
                            as.character(round(icc_PREDcog_DZS11$lbound,2)),"-",
                            as.character(round(icc_PREDcog_DZS11$ubound,2)),"]")
  
  fishZ_table[6,2] <- paste(as.character(round(icc_EMPcog_DZS11$value,2)),"[",
                            as.character(round(icc_EMPcog_DZS11$lbound,2)),"-",
                            as.character(round(icc_EMPcog_DZS11$ubound,2)),"]")
  
  fishZ_table[7,2] <- paste(as.character(round(icc_PREDcog_DZS13$value,2)),"[",
                            as.character(round(icc_PREDcog_DZS13$lbound,2)),"-",
                            as.character(round(icc_PREDcog_DZS13$ubound,2)),"]")
  
  fishZ_table[8,2] <- paste(as.character(round(icc_EMPcog_DZS13$value,2)),"[",
                            as.character(round(icc_EMPcog_DZS13$lbound,2)),"-",
                            as.character(round(icc_EMPcog_DZS13$ubound,2)),"]")
  
  fishZ_table[1,3] <- paste(as.character(round(icc_PREDemo_DZO11$value,2)),"[",
                            as.character(round(icc_PREDemo_DZO11$lbound,2)),"-",
                            as.character(round(icc_PREDemo_DZO11$ubound,2)),"]") 
  
  fishZ_table[2,3] <- paste(as.character(round(icc_EMPemo_DZO11$value,2)),"[",
                            as.character(round(icc_EMPemo_DZO11$lbound,2)),"-",
                            as.character(round(icc_EMPemo_DZO11$ubound,2)),"]")
  
  fishZ_table[3,3] <- paste(as.character(round(icc_PREDemo_DZO13$value,2)),"[",
                            as.character(round(icc_PREDemo_DZO13$lbound,2)),"-",
                            as.character(round(icc_PREDemo_DZO13$ubound,2)),"]") 
  
  fishZ_table[4,3] <- paste(as.character(round(icc_EMPemo_DZO13$value,2)),"[",
                            as.character(round(icc_EMPemo_DZO13$lbound,2)),"-",
                            as.character(round(icc_EMPemo_DZO13$ubound,2)),"]")
  
  
  fishZ_table[5,3] <- paste(as.character(round(icc_PREDcog_DZO11$value,2)),"[",
                            as.character(round(icc_PREDcog_DZO11$lbound,2)),"-",
                            as.character(round(icc_PREDcog_DZO11$ubound,2)),"]")
  
  fishZ_table[6,3] <- paste(as.character(round(icc_EMPcog_DZO11$value,2)),"[",
                            as.character(round(icc_EMPcog_DZO11$lbound,2)),"-",
                            as.character(round(icc_EMPcog_DZO11$ubound,2)),"]")
  
  fishZ_table[7,3] <- paste(as.character(round(icc_PREDcog_DZO13$value,2)),"[",
                            as.character(round(icc_PREDcog_DZO13$lbound,2)),"-",
                            as.character(round(icc_PREDcog_DZO13$ubound,2)),"]")
  
  fishZ_table[8,3] <- paste(as.character(round(icc_EMPcog_DZO13$value,2)),"[",
                            as.character(round(icc_EMPcog_DZO13$lbound,2)),"-",
                            as.character(round(icc_EMPcog_DZO13$ubound,2)),"]")
  
  
  fishZ_table[1,4] <- round(fishZ_predEmo11$z,2)
  fishZ_table[2,4] <- round(fishZ_empEmo11$z,2)
  fishZ_table[3,4] <- round(fishZ_predEmo13$z,2) 
  fishZ_table[4,4] <- round(fishZ_empEmo13$z,2)
  fishZ_table[5,4] <- round(fishZ_predCog11$z,2)
  fishZ_table[6,4] <- round(fishZ_empCog11$z,2)
  fishZ_table[7,4] <- round(fishZ_predCog13$z,2) 
  fishZ_table[8,4] <- round(fishZ_empCog13$z,2)

  colnames(fishZ_table) <- c("variable","DZ-same sex", "DZ-opposite sex","Z value")
  write.csv(fishZ_table, file= "fisherZ_DZ.csv", row.names = F)

```



uniting the DZS and DZO groups
```{r}
  DWide$zygosity_2G <- DWide$zygosity
  DWide$zygosity_2G[DWide$zygosity==1] <-"MZ"
  DWide$zygosity_2G[DWide$zygosity==2] <-"DZ"
  DWide$zygosity_2G[DWide$zygosity==3] <-"DZ"
```


compute ICC of the new variables separately for MZ and DZ twins (residualized from sex)
```{r}
  col <- colnames(DWide)

  icc_PREDemo_MZ11r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="predicted_emotional11r_1"),which(col=="predicted_emotional11r_2"))],model="oneway", unit="single")
  icc_PREDemo_DZ11r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="predicted_emotional11r_1"),which(col=="predicted_emotional11r_2"))],model="oneway", unit="single")
  
  
  icc_EMPemo_MZ11r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="EMPQ_emotional11r_1"),which(col=="EMPQ_emotional11r_2"))],model="oneway", unit="single")
  icc_EMPemo_DZ11r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="EMPQ_emotional11r_1"),which(col=="EMPQ_emotional11r_2"))],model="oneway", unit="single")
  
  icc_PREDemo_MZ13r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="predicted_emotional13r_1"),which(col=="predicted_emotional13r_2"))],model="oneway", unit="single")
  icc_PREDemo_DZ13r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="predicted_emotional13r_1"),which(col=="predicted_emotional13r_2"))],model="oneway", unit="single")
  
  
  icc_EMPemo_MZ13r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="EMPQ_emotional13r_1"),which(col=="EMPQ_emotional13r_2"))],model="oneway", unit="single")
  icc_EMPemo_DZ13r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="EMPQ_emotional13r_1"),which(col=="EMPQ_emotional13r_2"))],model="oneway", unit="single")
  
  
  icc_PREDcog_MZ11r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="predicted_cognitive11r_1"),which(col=="predicted_cognitive11r_2"))],model="oneway", unit="single")
  icc_PREDcog_DZ11r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="predicted_cognitive11r_1"),which(col=="predicted_cognitive11r_2"))],model="oneway", unit="single")
  
  
  icc_EMPcog_MZ11r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="EMPQ_cognitive11r_1"),which(col=="EMPQ_cognitive11r_2"))],model="oneway", unit="single")
  icc_EMPcog_DZ11r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="EMPQ_cognitive11r_1"),which(col=="EMPQ_cognitive11r_2"))],model="oneway", unit="single")
  
  icc_PREDcog_MZ13r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="predicted_cognitive13r_1"),which(col=="predicted_cognitive13r_2"))],model="oneway", unit="single")
  icc_PREDcog_DZ13r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="predicted_cognitive13r_1"),which(col=="predicted_cognitive13r_2"))],model="oneway", unit="single")
  
  icc_EMPcog_MZ13r  <- icc(DWide[DWide$zygosity_2G=="MZ",c(which(col=="EMPQ_cognitive13r_1"),which(col=="EMPQ_cognitive13r_2"))],model="oneway", unit="single")
  icc_EMPcog_DZ13r  <- icc(DWide[DWide$zygosity_2G=="DZ",c(which(col=="EMPQ_cognitive13r_1"),which(col=="EMPQ_cognitive13r_2"))],model="oneway", unit="single")


```

prepare a file only with the final variables for Mx biometric cross-lagged analysis
*notice that before entering the file to the Mx software, you need to erase the columns names 
 and to turn NA values into 999 (this was done manually in the .csv file)
```{r}
  col <- colnames(DWide)
  relvar_4mx <- c(which (col=="ifam"),which(col=="zygosity_2G"),which(col=="sex_2"),which(col=="sex_1"),
                  which(col=="predicted_emotional11r_2"),which(col=="predicted_emotional13r_2"),
                  which(col=="EMPQ_emotional11r_2"),which(col=="EMPQ_emotional13r_2"),
                  which(col=="predicted_cognitive11r_2"),which(col=="predicted_cognitive13r_2"),
                  which(col=="EMPQ_cognitive11r_2"),which(col=="EMPQ_cognitive13r_2"),
                  which(col=="predicted_emotional11r_1"),which(col=="predicted_emotional13r_1"),
                  which(col=="EMPQ_emotional11r_1"),which(col=="EMPQ_emotional13r_1"),
                  which(col=="predicted_cognitive11r_1"),which(col=="predicted_cognitive13r_1"),
                  which(col=="EMPQ_cognitive11r_1"),which(col=="EMPQ_cognitive13r_1"))
  
  DWide_4Mx <- DWide[,relvar_4mx]
  write.csv(DWide_4Mx, file= "DWide_4Mx.csv", row.names = F)         
```


Check the associations between the study variables and non-self-report empathy measures
```{r}

#import the data
  #Parent-reported empathy age 11
  DParentEmp <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Age11_ParentEmp_anonymized.csv")

  #erase invalid observations (e.g., mother rated instead of the father, parents completed the questionnaire together, etc.)
  DParentEmp <- DParentEmp[DParentEmp$whichParentOut!=3,]
  #erase father's data if only the father's data was invalid
  col <- colnames(DParentEmp)
  fathervars <- which(col=="EQSQ_empathy.2")
  DParentEmp[DParentEmp$whichParentOut==2,fathervars] <- NA

  #merge datasets
  relvar_parents <- c(which (col=="ifam"),which(col=="ID"),
                      which (col=="EQSQ_empathy.2"),which(col=="EQSQ_empathy.3"))
  DParentEmpShort <- DParentEmp[,relvar_parents]
  D<- merge(x=D, y=DParentEmpShort, by=c("ifam","ID"), all.x=T, all.y=T )  
  

  #Parent-reported empathy age 13
  D13ParentEmp <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Age13_ParentEmp_anonymized.csv")
  #erase invalid observations (e.g., mother rated instead of the father, parents completed the questionnaire together, etc.)
  D13ParentEmp <- D13ParentEmp[D13ParentEmp$whichParentOut==0,]
  
  col <- colnames(D13ParentEmp)
  relvar_parents13 <- c(which (col=="ifam"),which(col=="ID"),
                        which (col=="EQSQ_empathy.2"),which(col=="EQSQ_empathy.3"))
  D13ParentEmpShort <- D13ParentEmp[,relvar_parents13]

  #rename age 13 variables (enter indication of age 13 time-point)
  D13ParentEmpShort <- rename(D13ParentEmpShort, 
                            c(EQSQ_empathy.2="EQSQ_empathy_13.2",EQSQ_empathy.3="EQSQ_empathy_13.3"))
  D<- merge(x=D, y=D13ParentEmpShort, by=c("ifam","ID"), all.x=T, all.y=T )              

  
  #subtle facial expressions emotion recognition
  DEmoR <- read.csv ("~/Documents/projects/Empathy-Personality-Adolecence/OSF/data/Age13_emotion_recognition_test_finalvars_anonymized.csv")
  #remove children who did not have data from all emotion categories (see details in Appendix 1)
  DEmoRshort <- DEmoR[complete.cases(DEmoR[which(colnames(DEmoR)=="ER_c")]),]

  D <- merge(x=D, y=DEmoRshort, by=c("ifam","ID"), all.x=T, all.y=T ) 


#check difference in study variables according to parental missingness 
#(difference between children who had no parent data and children who had at least one parent data)
  D$missmomques11 <- NA
  D$missmomques11[!is.na(D$EQSQ_empathy.3)] <-0
  D$missmomques11[is.na(D$EQSQ_empathy.3)] <-1
  
  D$missdadques11 <- NA
  D$missdadques11[!is.na(D$EQSQ_empathy.2)] <-0
  D$missdadques11[is.na(D$EQSQ_empathy.2)] <-1
  
  D$missmomques13 <- NA
  D$missmomques13[!is.na(D$EQSQ_empathy_13.3)] <-0
  D$missmomques13[is.na(D$EQSQ_empathy_13.3)] <-1
  
  D$missdadques13 <- NA
  D$missdadques13[!is.na(D$EQSQ_empathy_13.2)] <-0
  D$missdadques13[is.na(D$EQSQ_empathy_13.2)] <-1
  
  D$missparentdata11 <- 0
  D$missparentdata11[D$missmomques11==1 & D$missdadques11] <-1
  D$missparentdata13 <- 0
  D$missparentdata13[D$missmomques13==1 & D$missdadques13] <-1
  
  #t-tests
  t.test(predicted_emotional_11~missparentdata11, data=D)
  t.test(predicted_cognitive_11~missparentdata11, data=D)
  t.test(EMPQ_emotional_11~missparentdata11, data=D)
  t.test(EMPQ_cognitive_11~missparentdata11, data=D)
  
  t.test(predicted_emotional_13~missparentdata13, data=D)
  t.test(predicted_cognitive_13~missparentdata13, data=D)
  t.test(EMPQ_emotional_13~missparentdata13, data=D)
  t.test(EMPQ_cognitive_13~missparentdata13, data=D)
  
#examine the correlations between study variables and non-self report measures
  col <- colnames(D)
  relvar <- c(which (col=="predicted_emotional_11"),which(col=="predicted_cognitive_11"),
            which (col=="predicted_emotional_13"),which(col=="predicted_cognitive_13"),
            which (col=="EMPQ_emotional_11"),which(col=="EMPQ_cognitive_11"),
            which (col=="EMPQ_emotional_13"),which(col=="EMPQ_cognitive_13"),
            which (col=="EQSQ_empathy.2"),which(col=="EQSQ_empathy.3"),
            which (col=="EQSQ_empathy_13.2"),which(col=="EQSQ_empathy_13.3"),
            which (col=="ER_c"), which (col=="ER_c_neg"),which (col=="ER_c_pos"))

  apa.cor.table(D[,relvar], filename = "EmpcorTable.doc")
  
```

